---
title: "Missing Spending Data Report"
author: "Dirk Cotton"
date: "4/17/2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
 load ("saveCams")
 knitr::opts_chunk$set(echo = TRUE)
 hrs <- read.csv("HRS Retired.csv")
 retiredHHID <- hrs$hhid[hrs$selectedBy != "notRetired"]
```

We identify `r sum(hrs$selectedBy != "notRetired")` retired households from the HRS database. However, spending data in the CAMS database *for these retired households* is available for only `r sum(cams$hhid %in% retiredHHID)` of those. To replace the missing spending data in our data set of retired households identified from HRS data, we impute retired household spending when the data is missing from CAMS.

CAMS household data is available for `r retiredHRSinCams` of the households identified as retired in HRS data.

We are interested in household spending for the wave in which each household was determined to have first retired (wave of retirement, or WOR). CAMS data is available beginning with wave 5 so we first remove all CAMS households determined from HRS data to have retired prior to wave 5 or not to have retired. `r beforeW5` households are removed from our CAMS data set on this basis, leaving `r afterW5 `households.

`r noWOR` of the remaining CAMS households provide no spending data for the wave of retirement. These observations are rejected leaving `r hasWOR` households in our CAMS data set.

We then reject `r noIncomeSave` CAMS households that have neither data for current income or any savings in HRS, despite providing consumption data, as we cannot identify the source of this consumption. This action leaves `r afterNoInc` observations in the CAMS data set.

These actions leave `r dim(cams1P)[1]` one-person households and `r dim(cams2P)[1]` two-person households with which we can model missing non-durable consumption expenditures by household when that data is missing from CAMS.

We impute spending separately for one-person and two-person households using multiple linear regression models.

For one-person households, the y-variable for our model is consumer non-durable expenditures for the household for the wave of retirement. The model's x-variables are:

  + Social Security income for the wave of retirement
  + Total portfolio assets for the wave of retirement
  + Pension and annuity income for the wave of retirement
  + age at retirement
  + gender (equals 1 for male, else zero), and
  + Wave number of retirement (5 through 12)

Wave number is modeled as a series of dummy variables, d5 through d12. If a household retired in wave 11, for example, d11 for that household will equal 1 and dummy variables for all other waves will equal zero for that household. (Only one of the eight dummy variables for a household will equal 1.)

The multiple progression is performed suppressing calculation of a y-intercept value. All x-variables expressed in dollars are normalized by dividing each x-variable for a household by the sum of all dollar-denominated x-variables for that household.

The results of the linear regression analysis for one-person households are summarized:

```{r warnings=FALSE, echo=FALSE}
summary(fit)
```

A few modifications are made to the regression model for two-person households:

  - All HRS two-person households consist of one male and one female, so gender is not a factor
  - Social Security retirement income for the second household member must be added to the model
  - The age of the second household member must also be added to the model
  
The results of the multiple regression analysis for two-person households are summarized:

```{r warnings=FALSE, echo=FALSE}
summary(fit2)
```


Coefficients determined by the respective linear regression results are used to impute a "predicted" amount of consumer non-durable spending for retired households from the HRS data base when no spending is available for the household's wave of retirement.

