---
title: "Report Cross Validation of Bad Error Predictions"
author: "Dirk Cotton"
date: "5/9/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load ("~/R Projects/HRS/CrossVal.RData")
```

The consumption data predicted by our models based on HRS wealth and income data showed large errors relative to reported consumption from the CAMS data base for a large number of households. We cross-validated our models to determine if removing those households with poor predictions would improve prediction results.

We us a model for one-person households. The regression model for one-person households follows:

```{r warning=FALSE,echo=FALSE}
print(fit)

```

We predicted consumption using the "fit" model of predictions with acceptable errors (< 100%), randomly selecting `r length(highError1Phhid)` households with acceptable errors as the prediction set from a data set of `r dim(cams1Pgtw)[1]` CAMS households.

For the second cross-validation, we used `r length(highError1Phhid)` households with high errors as the predictions set.

We plotted the density of the eroor results (predicted consumption minus reported consumption) and generated the following two plots:

```{r warning=FALSE,echo=FALSE}

source('~/R Projects/HRS/CrossValidate 1P HH Good Predictions.R')

plot(density(crossValOnePerson[,1]),main="1-Person Household Good Predictions\nPredicted - Actual",ylim=c(0,.0001),xlab="Prediction Error ($)")
for (i in 2:n){
  lines(density(crossValOnePerson[,i]))
}

```


```{r warning=FALSE,echo=FALSE}
source('~/R Projects/HRS/CrossValidate 1P HH Good Predictions.R')
plot(density(crossValOnePerson[,1]),main="1-Person Household Bad Predictions\nPredicted - Actual",xlab="Prediction Error ($)")
for (i in 2:n){
  lines(density(crossValOnePerson[,i]))
}
```

Our conclusion from these results is that removing households with poor predictions from the model will not improve the model's predictions.


NOTES: 

[1] This output is created by running the following scripts in order:

1. source('~/R Projects/HRS/Validate Predicted CAMS Data.R')
2. source('~/R Projects/HRS/CrossValidate 1P HH Good Predictions.R')
3. source('~/R Projects/HRS/CrossValidate 1P HH Bad Predictions.R')

and then knitting the Word document with this script.

[2] The errant line on the "Good Predictions" graph is caused by a household with an extreme negative prediction that doesn't fit within the graph's scale.
